---
title: "Data Analysis Process with SQL"
output:
  html_document:
    toc: true
    toc_float: true
    collapsed: false
    number_sections: false
    toc_depth: 1
    #code_folding: hide
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(message=FALSE,warning=FALSE, cache=TRUE)
```
<br>

# Explanation
<hr>
Below I explain all the steps that I took in SQL for my data analysis process. I believe it's important to do this case study with both SQL and R. The idea is to become more proficient in both. 

Before you go further with this page, I strongly recommend you to read my <a href="https://ru4871sg.github.io/Google-Data-Analytics-Professional-Cyclistic-Capstone/process.html">data analysis process in R</a>. Over there, I have much better explanation for everything.

Here you can find the alternative steps in SQL, however, I won't write much details here because I don't want to keep repeating myself to explain the same steps.

*Note: My project folder in BigQuery is called `coursera-test-367602`. That's why you keep seeing this project folder in the below steps.

<br><br><br>

# 1. Import Datasets to BigQuery
<hr>
I can't import half of the datasets directly. Datasets from May 2022 to October 2022 are larger than 100 MB. I have to convert these large files to Excel Workbook files, and then, I need to make sure to use `yyyy-mm-dd h:mm:ss` format for some of the columns.<br><br>

# 2. Slice Larger Files to Two Different Sheets 
<hr>
By slicing them with Microsoft Excel, the file sizes will be smaller (this is needed to bypass the 100 MB limitation in BigQuery). For example, <b>may_1.csv</b> and <b>may_2.csv</b>, <b>june_1.csv</b> and <b>june_2.csv</b>, and so on. The rest of csv which originally are less than 100 MB, I don't need to slice them. Import all to BigQuery.


# 3. Use UNION ALL for All Imported Tables
<hr>
```{r, eval=F, echo=T}
SELECT *
FROM `coursera-test-367602.cyclistic_case_study.cyclistic_tripdata_202201`
UNION ALL 
SELECT *
FROM `coursera-test-367602.cyclistic_case_study.cyclistic_tripdata_202202`
UNION ALL 
SELECT *
FROM `coursera-test-367602.cyclistic_case_study.cyclistic_tripdata_202203`
UNION ALL 
SELECT *
FROM `coursera-test-367602.cyclistic_case_study.cyclistic_tripdata_202204`
UNION ALL 
SELECT *
FROM `coursera-test-367602.cyclistic_case_study.cyclistic_tripdata_202205`
UNION ALL 
SELECT *
FROM `coursera-test-367602.cyclistic_case_study.cyclistic_tripdata_202206`
UNION ALL 
SELECT *
FROM `coursera-test-367602.cyclistic_case_study.cyclistic_tripdata_202207`
UNION ALL 
SELECT *
FROM `coursera-test-367602.cyclistic_case_study.cyclistic_tripdata_202208`
UNION ALL 
SELECT *
FROM `coursera-test-367602.cyclistic_case_study.cyclistic_tripdata_202209`
UNION ALL 
SELECT *
FROM `coursera-test-367602.cyclistic_case_study.cyclistic_tripdata_202210`
UNION ALL 
SELECT *
FROM `coursera-test-367602.cyclistic_case_study.cyclistic_tripdata_202211`
UNION ALL 
SELECT *
FROM `coursera-test-367602.cyclistic_case_study.cyclistic_tripdata_202212`
```
<br>
This will stack all the sheets on top of each other into one year. It works like `bind_rows()` in R, but `UNION ALL` is in SQL.<br><br>

# 4. Sort Table
<hr>
```{r, eval=F, echo=T}
SELECT * FROM `coursera-test-367602.cyclistic_case_study.cyclistic_2022_full`
ORDER BY started_at
```
<br><br>

# 5. Calculate ride_length and day_of_week
<hr>
```{r, eval=F, echo=T}
SELECT *,
    FORMAT("%02d:%02d:%02d",
        EXTRACT(HOUR FROM ride_length_preformatted),
        EXTRACT(MINUTE FROM ride_length_preformatted),
        EXTRACT(SECOND FROM ride_length_preformatted)) AS ride_length,
        EXTRACT(HOUR FROM ride_length_preformatted) AS hour,
        EXTRACT(MINUTE FROM ride_length_preformatted) AS minute,
        EXTRACT(SECOND FROM ride_length_preformatted) AS second,
    EXTRACT(DAYOFWEEK FROM started_at) AS day_of_week
FROM (
  SELECT *,
       ended_at - started_at AS ride_length_preformatted
  FROM `coursera-test-367602.cyclistic_case_study.cyclistic_2022_full_sorted`
)
AS add_ride_length
ORDER BY ride_length
```
<br>
Save it to `cyclistic_2022_precleaned`<br><br>

# 6. Premature Data Cleaning
<hr>
```{r, eval=F, echo=T}
SELECT *
FROM (SELECT *, 
      (hour * 3600) + (minute * 60) + second AS total_seconds
      FROM `coursera-test-367602.cyclistic_case_study.cyclistic_2022_precleaned` 
)
  AS with_total_seconds
WHERE total_seconds >= 60
ORDER BY total_seconds
```
<br>
Once there is no negative ride_length anymore, filter out `NULL` values as well, the main query ends up like this:
```{r, eval=F, echo=T}
SELECT * 
FROM (SELECT *, 
      (hour * 3600) + (minute * 60) + second AS total_seconds
      FROM `coursera-test-367602.cyclistic_case_study.cyclistic_2022_precleaned` 
)
  AS with_total_seconds
WHERE total_seconds >= 60
  AND minute > 0
  AND start_station_name IS NOT NULL
  AND start_station_id IS NOT NULL
  AND end_station_name IS NOT NULL
  AND end_station_id IS NOT NULL
  AND ride_id IS NOT NULL
  AND rideable_type IS NOT NULL
  AND started_at IS NOT NULL
  AND ended_at IS NOT NULL
  AND start_lat IS NOT NULL
  AND start_lng IS NOT NULL
  AND end_lat IS NOT NULL
  AND end_lng IS NOT NULL
  AND member_casual IS NOT NULL
```
<br><br>

# 7. Check Other Possibilities for Further Data Cleaning
<hr>
Next, check other possibilities with the second query tab:
```{r, eval=F, echo=T}
SELECT * 
FROM `coursera-test-367602.cyclistic_case_study.cyclistic_2022_precleaned` 
WHERE UPPER(start_station_name) = start_station_name
```
<br>
Let's do the same for `end_station_name`:
```{r, eval=F, echo=T}
SELECT * 
FROM `coursera-test-367602.cyclistic_case_study.cyclistic_2022_precleaned` 
WHERE UPPER(end_station_name) = end_station_name
```
<br>
Then, let's check if any row has the word "test" or "Test" or "TEST", who knows there are test related trips.
```{r, eval=F, echo=T}
SELECT * 
FROM `coursera-test-367602.cyclistic_case_study.cyclistic_2022_precleaned` 
WHERE start_station_name LIKE "%TEST%"
  OR start_station_name LIKE "%test%"
  OR start_station_name LIKE "%Test%"
  OR start_station_id LIKE "%TEST%"
  OR start_station_id LIKE "%test%"
  OR start_station_id LIKE "%Test%"
  OR end_station_name LIKE "%TEST%"
  OR end_station_name LIKE "%test%"
  OR end_station_name LIKE "%Test%"
  OR end_station_id LIKE "%TEST%"
  OR end_station_id LIKE "%test%"
  OR end_station_id LIKE "%Test%"
```
<br>
Just like how I've explained it in the data analysis process with R page, this step is very important to find out what types of row values that need to be filtered.
<br><br>

# 8. Final Data Cleaning
<hr>
```{r, eval=F, echo=T}	
UPDATE `coursera-test-367602.cyclistic_case_study.cyclistic_2022_precleaned` 
SET start_station_name = REPLACE(start_station_name, '*', '')

UPDATE `coursera-test-367602.cyclistic_case_study.cyclistic_2022_precleaned` 
SET start_station_name = REPLACE(start_station_name, '(Temp)', '')

UPDATE `coursera-test-367602.cyclistic_case_study.cyclistic_2022_precleaned` 
SET start_station_name = REPLACE(start_station_name, '- Charging', '')


UPDATE `coursera-test-367602.cyclistic_case_study.cyclistic_2022_precleaned` 
SET end_station_name = REPLACE(end_station_name, '*', '')

UPDATE `coursera-test-367602.cyclistic_case_study.cyclistic_2022_precleaned` 
SET end_station_name = REPLACE(end_station_name, '(Temp)', '')

UPDATE `coursera-test-367602.cyclistic_case_study.cyclistic_2022_precleaned` 
SET end_station_name = REPLACE(end_station_name, '- Charging', '')
```	
<br>
Once the table is updated, let's use this:
```{r, eval=F, echo=T}	
SELECT DISTINCT ride_id, rideable_type, started_at, ended_at, TRIM(start_station_name) AS start_station_name, TRIM(end_station_name) AS end_station_name, start_station_id, end_station_id, start_lat, start_lng, end_lat, end_lng, TRIM(member_casual) AS member_casual, ride_length, total_seconds, day_of_week 
FROM (SELECT *, 
      (hour * 3600) + (minute * 60) + second AS total_seconds
      FROM `coursera-test-367602.cyclistic_case_study.cyclistic_2022_precleaned` 
)
  AS with_total_seconds

WHERE total_seconds >=60 
  AND start_station_name IS NOT NULL
  AND start_station_id IS NOT NULL
  AND end_station_name IS NOT NULL
  AND end_station_id IS NOT NULL
  AND ride_id IS NOT NULL
  AND rideable_type IS NOT NULL
  AND started_at IS NOT NULL
  AND ended_at IS NOT NULL
  AND start_lat IS NOT NULL
  AND start_lng IS NOT NULL
  AND end_lat IS NOT NULL
  AND end_lng IS NOT NULL
  AND member_casual IS NOT NULL
  AND UPPER(start_station_name) != start_station_name
  AND UPPER(end_station_name) != end_station_name
  AND start_station_name NOT LIKE "%TEST%"
  AND start_station_name NOT LIKE "%test%"
  AND start_station_name NOT LIKE "%Test%"
  AND start_station_id NOT LIKE "%TEST%"
  AND start_station_id NOT LIKE "%test%"
  AND start_station_id NOT LIKE "%Test%"
  AND end_station_name NOT LIKE "%TEST%"
  AND end_station_name NOT LIKE "%test%"
  AND end_station_name NOT LIKE "%Test%"
  AND end_station_id NOT LIKE "%TEST%"
  AND end_station_id NOT LIKE "%test%"
  AND end_station_id NOT LIKE "%Test%"
ORDER BY started_at
```	
<br>
Save the result to `cyclistic_2022_postcleaned`<br><br>

# 9. Analyze Most Popular Station Count
<hr>
### 9a. Most Popular Stations for Members
let's start with `start_station_name`:
```{r, eval=F, echo=T}
SELECT start_station_name AS station, COUNT(start_station_name) AS start_station_name_count
FROM `coursera-test-367602.cyclistic_case_study.cyclistic_2022_postcleaned` 
WHERE member_casual = 'member'
GROUP BY start_station_name
ORDER BY start_station_name_count DESC
```
<br>
Now, let's do the same for `end_station_name`:
```{r, eval=F, echo=T}
SELECT end_station_name AS station, COUNT(end_station_name) AS end_station_name_count
FROM `coursera-test-367602.cyclistic_case_study.cyclistic_2022_postcleaned` 
WHERE member_casual = 'member'
GROUP BY end_station_name
ORDER BY end_station_name_count DESC
```
<br>
Save both of the above results as new tables, save them as `start_station_name_count_member` and `end_station_name_count_member`. Next, let's use `FULL OUTER JOIN`:
```{r, eval=F, echo=T}
SELECT 
  CASE WHEN start_station.station IS NOT NULL THEN start_station.station
  ELSE end_station.station
  END AS station_name_both,
  
CASE WHEN start_station_name_count IS NOT NULL AND end_station_name_count IS NOT NULL THEN start_station_name_count + end_station_name_count
  WHEN start_station_name_count IS NULL THEN end_station_name_count
  ELSE start_station_name_count
  END AS station_count_both
FROM 
  `coursera-test-367602.cyclistic_case_study.start_station_name_count_member` AS start_station -- COMMENT: TABLE 1
FULL OUTER JOIN
  `coursera-test-367602.cyclistic_case_study.end_station_name_count_member` AS end_station -- COMMENT: TABLE 2
  ON start_station.station = end_station.station 
ORDER BY station_count_both DESC
```
I save it as `count_station_name_member`
<br><br>

### 9b. Most Popular Stations for Casuals
let's start with `start_station_name`:
```{r, eval=F, echo=T}
SELECT start_station_name AS station, COUNT(start_station_name) AS start_station_name_count
FROM `coursera-test-367602.cyclistic_case_study.cyclistic_2022_postcleaned` 
WHERE member_casual = 'casual'
GROUP BY start_station_name
ORDER BY start_station_name_count DESC
```
<br>
Next, let's do the same for `end_station_name`:
```{r, eval=F, echo=T}
SELECT end_station_name AS station, COUNT(end_station_name) AS end_station_name_count
FROM `coursera-test-367602.cyclistic_case_study.cyclistic_2022_postcleaned` 
WHERE member_casual = 'casual'
GROUP BY end_station_name
ORDER BY end_station_name_count DESC
```
<br>
Save both of the above results as new tables, `start_station_name_count_casual` and `end_station_name_count_casual`. And just like before, use `FULL OUTER JOIN`:
```{r, eval=F, echo=T}
SELECT 
  CASE WHEN start_station.station IS NOT NULL THEN start_station.station
  ELSE end_station.station
  END AS station_name_both,
  
CASE WHEN start_station_name_count IS NOT NULL AND end_station_name_count IS NOT NULL THEN start_station_name_count + end_station_name_count
  WHEN start_station_name_count IS NULL THEN end_station_name_count
  ELSE start_station_name_count
  END AS station_count_both
FROM 
  `coursera-test-367602.cyclistic_case_study.start_station_name_count_casual` AS start_station -- COMMENT: TABLE 1
FULL OUTER JOIN
  `coursera-test-367602.cyclistic_case_study.end_station_name_count_casual` AS end_station -- COMMENT: TABLE 2
  ON start_station.station = end_station.station 
ORDER BY station_count_both DESC
```
<br>
I save it as `count_station_name_casual`
<br><br>

# 10. Analyze Popular Days of Week and Dates
<hr>
```{r, eval=F, echo=T}
SELECT day_of_week, member_casual, COUNT(day_of_week) AS day_of_week_count 
FROM `coursera-test-367602.cyclistic_case_study.cyclistic_2022_postcleaned` 
GROUP BY day_of_week, member_casual
ORDER BY day_of_week_count DESC
```
<br>
Save this result as `count_day_of_week`. Next:
```{r, eval=F, echo=T}
SELECT day_of_week, member_casual, month, COUNT(month) AS month_count
FROM (
SELECT *,
      EXTRACT (MONTH FROM started_at) AS month
FROM `coursera-test-367602.cyclistic_case_study.cyclistic_2022_postcleaned` 
WHERE member_casual = 'casual'
)
  AS with_clean_date
GROUP BY month, day_of_week, member_casual
ORDER BY month_count DESC
```
<br>
and
```{r, eval=F, echo=T}
SELECT day_of_week, member_casual, month, COUNT(month) AS month_count
FROM (
SELECT *,
      EXTRACT (MONTH FROM started_at) AS month
FROM `coursera-test-367602.cyclistic_case_study.cyclistic_2022_postcleaned` 
WHERE member_casual = 'member'
)
  AS with_clean_date
GROUP BY month, day_of_week, member_casual
ORDER BY month_count DESC
```
<br>
Save both results, as `count_popular_dates_casual` and `count_popular_dates_member`.<br><br>

# 11. Analyze Most Popular Hours of the Day
<hr>
Let's start with total count for all (both casual and member):
```{r, eval=F, echo=T}
SELECT clean_hour, COUNT(clean_hour) AS clean_hour_count
FROM (
SELECT *, 
    FORMAT("%02d",
        EXTRACT(HOUR FROM started_at)) AS clean_hour
FROM `coursera-test-367602.cyclistic_case_study.cyclistic_2022_postcleaned` 
)
  AS with_clean_hour
GROUP BY clean_hour
ORDER BY clean_hour_count DESC
```
<br>
Next, let's divide them by the casual or member groups.
```{r, eval=F, echo=T}
SELECT clean_hour, COUNT(clean_hour) AS clean_hour_count, member_casual
FROM (
SELECT *, 
    FORMAT("%02d",
        EXTRACT(HOUR FROM started_at)) AS clean_hour
FROM `coursera-test-367602.cyclistic_case_study.cyclistic_2022_postcleaned` 
WHERE member_casual = 'member' 
)
  AS with_clean_hour
GROUP BY clean_hour, member_casual
ORDER BY clean_hour_count DESC
```
<br>
and
```{r, eval=F, echo=T}
SELECT clean_hour, COUNT(clean_hour) AS clean_hour_count, member_casual
FROM (
SELECT *, 
    FORMAT("%02d",
        EXTRACT(HOUR FROM started_at)) AS clean_hour
FROM `coursera-test-367602.cyclistic_case_study.cyclistic_2022_postcleaned` 
WHERE member_casual = 'casual' 
)
  AS with_clean_hour
GROUP BY clean_hour, member_casual
ORDER BY clean_hour_count DESC
```
<br>
Let's save both of the above queries as `count_popular_hours_casual` and `count_popular_hours_member`.<br><br>

# 12. Analyze Ride Length Difference Between Casual and Member
<hr>
```{r, eval=F, echo=T}
SELECT AVG(total_seconds / 60) AS avg_ride_length_in_minutes, member_casual 
FROM `coursera-test-367602.cyclistic_case_study.cyclistic_2022_postcleaned`
GROUP BY member_casual
ORDER BY avg_ride_length_in_minutes DESC
```
<br><br>

# 13. Analyze Bike Types
<hr>
Let's start with this:
```{r, eval=F, echo=T}
SELECT member_casual, rideable_type, COUNT(*) AS trips
FROM `coursera-test-367602.cyclistic_case_study.cyclistic_2022_postcleaned`
GROUP BY rideable_type, member_casual
ORDER BY trips DESC
```
<br>
Save this result as `count_rideable_type`. Next:
```{r, eval=F, echo=T}
SELECT member_casual, COUNT(*) AS total_trips
FROM `coursera-test-367602.cyclistic_case_study.cyclistic_2022_postcleaned`
GROUP BY member_casual
ORDER BY total_trips DESC
```
<br>
Save this result as `count_member_vs_casual`. Now let's use `JOIN`:
```{r, eval=F, echo=T}
SELECT rideable_type_count.member_casual AS member_casual,
       rideable_type_count.rideable_type AS rideable_type,
       rideable_type_count.trips AS trips_each_type,
       member_vs_casual.total_trips AS total_trips,
       (rideable_type_count.trips / member_vs_casual.total_trips) AS trips_percentage
FROM `coursera-test-367602.cyclistic_case_study.count_rideable_type` AS rideable_type_count
JOIN
  `coursera-test-367602.cyclistic_case_study.count_member_vs_casual` AS member_vs_casual
  ON rideable_type_count.member_casual = member_vs_casual.member_casual
```
<br>
Save this result as `count_and_percentage_rideable_type`. Now, let's split them by casual and member groups:
```{r, eval=F, echo=T}
SELECT member_casual, rideable_type, trips_percentage 
FROM `coursera-test-367602.cyclistic_case_study.count_and_percentage_rideable_type`
WHERE member_casual = 'casual' 
```
<br>
and
```{r, eval=F, echo=T}
SELECT member_casual, rideable_type, trips_percentage 
FROM `coursera-test-367602.cyclistic_case_study.count_and_percentage_rideable_type`
WHERE member_casual = 'member'
```
<br>
Save both results as `casual_rideable_type` and `member_rideable_type`. That's it. Export all the saved results and visualize them with R.